Comment: Inferring Trade Costs from Trade Booms and Trade Busts

### 2016-06-06 Paper/Manuscript.tex
Changement avec Esther

\author[1]{Guillaume Corlay}
\author[2]{Stéphane Dupraz}
\author[1]{Claire Labonne}
\author[1]{Anne Muller}
%\author[3]{Guillaume Daudin}
\author[3]{Guillaume Daudin\footnote{Corresponding author: guillaume.daudin@dauphine.fr, }\footnote{We are grateful to Celine Antonin (OFCE) for her constant supply of helpful comments. She, of course, bears no responsibility for remaining errors.}}
\affil[1] {ENSAE, French National School of Statistics and Economic Administration, 3, avenue Pierre Larousse, 92245 Malakoff, FRANCE}
\affil[2] {Columbia University, 10025 New York NY, USA}
\affil[3] {Université Paris-Dauphine, PSL Research University, LEDa, 75016 PARIS, FRANCE
Université Paris-Dauphine, PSL Research University, LEDa, UMR [225], DIAL, 75016 PARIS, FRANCE

@@

\begin{abstract}
\cite{JMN2011} offer an alternative to price gaps to quantify
trade costs. Their method consists in deducing international
trade costs from trade flows. They use it to calculate the
evolution of these costs during the first and second waves of
globalization (1870-1913 and 1950-2000) and the interwar period
(1921-1939). They reach the conclusion that the reduction in
trade costs was the main driving force of the growth of trade
during the first globalization, whereas the trade boom was led
by economic expansion during the second globalization. We argue
that this result is entirely driven by the ad hoc aggregation
method of trade costs the authors use. We show that what their
measure captures is instead a difference in the size of the
countries concerned by the phenomenon of globalization. More
generally, we cast doubts on the possibility to reach
conclusions of such nature with a method that infers trade
costs from trade flows, and then uses these costs to explain
trade flows, if not to rephrase the information already
contained in an openness ratio.
\end{abstract}

Keywords: Trade costs, globalization, gravity model, aggregation,
structure effect.
JEL Code:

@@

Jacks and his coauthors offer in a number of papers \footnote{The method is developed in\cite{JMN2008}, \cite{JMN2010}, \cite{NOVY} and \cite{JMN2011}. We will use this latter paper as a referenc} a method for
inferring international trade costs from trade flows. Using the general equilibrium model of
\cite{AW2003}, they calculate trade costs (defined as as all barriers to trade, notably transport
and transaction costs) and their
evolution during the first and second waves of globalization
(1870-1913 and 1950-2000) as well as the interwar period
(1921-1939) thanks to the impressive set of data they collected
on trade flows, GDP and exports between 27
countries\footnote{Argentina, Australia, Austria, Belgium,
Brazil, Canada, Denmark, France, Germany, Greece, India,
Indonesia, Italy, Japan, Mexico, the Netherlands, New Zealand,
Norway, the Philippines, Portugal, Spain, Sri Lanka, Sweden,
Switzerland, the United Kingdom, the United States, and
Uruguay. The data contain 130 country pairs.}. They provide a
decomposition of the growth of trade caused by the reduction in
trade costs and economic expansion. They use their computations to underline
a difference of nature between the two globalizations:

@@

We argue that this result is actually driven by an ad hoc
method of aggregation. When relying instead on a
three-countries version of \cite{AW2003}'s model for
aggregating bilateral trade costs - where the three countries
are the country of interest, its aggregate trade partners
present in the data, and the rest of the world - the alleged
difference of natures between the two globalizations fades
away. \cite{JMN2011}'s aggregation technique captures structure effects. The inferred difference between the
two globalization

????? comes from the evolving size of countris rather that trade costs ????

concerns the size of the countries comes from
the phenomenon of globalization, not trade costs. More generally, we
cast doubt on the possibility to distinguish the effect of trade costs and economic expansion through an approach that relies only on the study of trade
flows. Deducing trade costs from trade flows, and then using
trade costs to explain trade flows is essentially a circular
reasoning. Therefore, Jacks et al.'s approach cannot be an
alternative to traditional investigations of impediments to
trade, such as commodity price gaps.

@@

//

@@

The average of the bilateral trade costs $\tau_{ij}$ captures two elements: trade
costs and the distribution of these costs relative to the size
of the trading partners. Through time, the more small countries have a relatively faster growth of trade, the more the decline of trade costs measured by $\tau_i$ becomes is overestimated.
This fits the data. For both globalizations, figure
\ref{structure} plots the growth of trade flows (measured by
$\Delta ln (\sqrt{x_{ij} x_{ji}})$) as a function of the
initial value of trade, (measured by the logarithm of the
geometric average of bilateral average $ln(\sqrt{x_{ij}
x_{ji}})$) for all dyads in the sample. It confirms that the
growth of international trade in the first globalization is
concentrated on pairs of countries that initially traded little
together. This is also apparent in the contrast between the rate of growth of exports
(table \ref{OG}) and the rate of growth of bilateral trade flows (table \ref{glob1} and \ref{glob2}.
This is what misled Jacks, Meissner and Novy into
stating a difference of nature between the two globalizations.

@@

This characteristic of the proposed measure of trade costs is
partly due to the comprehensively inclusive definition of trade
costs used by the authors. In such a model, costs are anything
that causes consumptions from different countries' products not
to be equal. To be sure, it is essential to highlight that
trade costs cannot be reduced to tariffs or transportation
costs, and to insist on the need for a quantification of all
the impediments to international trade. But the concept of
trade costs loses part of its interest if there are no causal
alternatives to explain trade patterns. Such a definition is
therefore bound to always reword the information given by trade
flows relative to output, such as contained in an openness
ratio.

@@

\nocite{*}
\bibliographystyle{apalike}
\bibliography{bibliography}

>>>>

\author[1]{Guillaume Corlay}
\author[2]{Stéphane Dupraz}
\author[1]{Claire Labonne}
\author[1]{Anne Muller}
\author[3]{Guillaume Daudin\footnote{Corresponding author: guillaume.daudin@dauphine.fr, }\footnote{We are grateful to Celine Antonin (OFCE) for her constant supply of helpful comments. She, of course, bears no responsibility for remaining errors.}}
\affil[1] {ENSAE, French National School of Statistics and Economic Administration, 3, avenue Pierre Larousse, 92245 Malakoff, FRANCE}
\affil[2] {Columbia University, 10025 New York NY, USA}
\affil[3] {Université Paris-Dauphine, PSL Research University, LEDa, 75016 PARIS, FRANCE
Université Paris-Dauphine, PSL Research University, LEDa, UMR [225], DIAL, 75016 PARIS, FRANCE

@@

\begin{abstract}
\cite{JMN2011} offer an alternative to price gaps to quantify
trade costs. Their method consists in deducing international
trade costs from trade flows. They argue that the reduction in
trade costs was the main driving force of the growth of trade
during the first globalization (1870-1913), whereas economic expansion was the main driving force during the second globalization (1950-2000). We argue
that this important result is  driven by the use of an ad hoc aggregation
method. What \cite{JMN2011} capture is the 
difference in the relative starting trade of dyads experiencing faster trade growth in the first and second globalization.
More generally, we cast doubts on the possibility to reach
conclusions of such nature with a method that infers trade
costs from trade flows, and then uses these costs to explain
trade flows. It can only rephrase the information already
contained in openness ratios.
\end{abstract}

Keywords: Trade costs, globalization, gravity model, aggregation,
structure effect.

JEL Code: F14, N70

@@

Jacks and his coauthors offer in a number of papers  a method for
inferring international trade costs from trade flows.\footnote{The method is developed in\cite{JMN2008}, \cite{JMN2010}, \cite{NOVY} and \cite{JMN2011}. We will use this latter paper as a reference} Using the general equilibrium model of
\cite{AW2003}, they calculate trade costs (defined as as all barriers to trade, notably transport and transaction costs) and their
evolution during the first and second waves of globalization
(1870-1913 and 1950-2000) as well as the interwar period
(1921-1939) thanks to the impressive set of data they collected
on trade flows, GDP and exports between 27
countries\footnote{Argentina, Australia, Austria, Belgium,
Brazil, Canada, Denmark, France, Germany, Greece, India,
Indonesia, Italy, Japan, Mexico, the Netherlands, New Zealand,
Norway, the Philippines, Portugal, Spain, Sri Lanka, Sweden,
Switzerland, the United Kingdom, the United States, and
Uruguay. The data contain 130 country pairs.}. They provide a
decomposition of the growth of trade caused by the reduction in
trade costs and economic expansion. They use their computations to underline
a difference of nature between the two globalizations:

@@

This is potentially an important result that sheds light on the globalization processes. However, that result is actually driven by an ad hoc
method of aggregation that captures structure effects. The inferred difference between the two globalization comes from the fact that the dyads with the 
fastest growing trade in the first wave of globalization start with very small 
trade. This is not the case in the second wave of globalization.
Using a three-countries version of \cite{AW2003}'s model for
the aggregation of bilateral trade costs shows that there was no 
difference in nature betwee the two globalizations. More generally, we
cast doubt on the possibility to distinguish the effect of trade costs and economic expansion through an approach that relies only on the study of trade
flows. Deducing trade costs from trade flows, and then using
trade costs to explain trade flows is essentially a circular
reasoning. Therefore, Jacks et al.'s approach cannot be an
alternative to traditional investigations of impediments to
trade, such as commodity price gaps.

@@

@@

The average of the bilateral trade costs $\tau_{ij}$ captures both trade
costs and the distribution of trade costs relative to the size
of the trading partners. Through time, if dyads with small starting trade experience  faster growth of trade than others, the decline of trade costs measured by $\tau_i$ will be overestimated compared to the decline of $\upsilon_i$.
This fits the data. For both globalizations, figure
\ref{structure} plots the growth of trade flows (measured by
$\Delta ln (\sqrt{x_{ij} x_{ji}})$) as a function of the
initial value of trade, (measured by the logarithm of the
geometric average of bilateral average $ln(\sqrt{x_{ij}
x_{ji}})$) for all dyads in the sample. This is also confirmed in the contrast between the rate of growth of exports
(table \ref{OR}) and the rate of growth of bilateral trade flows (table \ref{glob1} and \ref{glob2}), as export growth was relatively faster than bilateral trade growth in the first globalization compared to the second (486/178 > 473/246). It confirms that pairs of countris that initially traded little together experienced faster growth of trade compared to other pairs in the first than in the second globalization.  This in an interesting result, but not the one presented by JMN.

@@

This characteristic of the proposed measure of trade costs is
partly due to the all-inclusive definition of trade
costs used by the authors. In such a model, costs are anything
that causes consumptions from different countries' products not
to be equal. To be sure, it is essential to highlight that
trade costs cannot be reduced to tariffs or transportation
costs, and to insist on the need for a quantification of all
the impediments to international trade. But the concept of
trade costs loses part of its interest if there are no causal
alternatives to explain trade patterns. Such a definition is
therefore bound to reword the information given by trade
flows relative to output, such as contained in an openness
ratio.

@@

\bibliographystyle{apalike}
\bibliography{bibliography}
